---
title: "creative-data-visualization"
author: "Jyhreh Johnson"
date: "5/12/2022"
output: html_document
---
```{r}
require(tidyverse)
require(dplyr)
require(ggplot2)
require(gganimate)
```
## Introduction
For this project, I looked attempted to create fun visualization of track and field (mainly events from the track) run time results for the Women & Men's 100M dashes. I played around with the idea of making the data race each other like actual track and field athlete do. **An ultimate goal is to graph an oval track and let the times race on the track**\

## Data Wrangling 
```{r}
f <- "https://raw.githubusercontent.com/jyhrehjohnson/creative-data-visualization/main/results.csv"
d <- read_csv(f, col_names = TRUE)
head(d)
```

## Data filtering 
```{r}
#select the data to remove location and athlete name
d <- d %>%
  select(Gender,
         Event,
         Year,
         Medal,
         Nationality,
         Result
         )
```

```{r} 
#filter event column for only Men and Women 100M dashes
d_100M_Men <- d %>% filter (Event == "100M Men")
d_100M_Men

d_100M_Women <- d %>% filter(Event == "100M Women")
d_100M_Women
```

## LET THE RACES BEGIN!!!!\

**Medal Times Race (Plot Race)**\
These two sets of visualizations, graph the Gold, Silver, & Bronze Medal times from the 1932 Olympics to the 2016 Olympics. 

Women's 100M Medal Times from 1932-2016
```{r}
w_Year_Tracker <- d_100M_Women %>% #pipe the filtered data through
  ggplot() + 
    geom_point(aes(x = Year, y = Result, col = Medal), 
    alpha = 0.8) + #set the parameter data for the plot
    theme(legend.position = "right") + #create & set position of legend
    guides(size = "none") + 
    labs(x = "Year",y = "Run Time Results (sec)", title = "Women's 100M Medal Times (1932 - 2016)", col = "Medal Type") + #set the x and y axis labels and title
    geom_text(aes(x = min(Year), y = min(Result), label = as.factor(Year)), hjust=0, vjust = -0.2, alpha = 0.2,  col = "gray", size = 20) + #use data to set column/row text and length of the graph
    transition_states(as.factor(Year), state_length = 0) #separates the data by year to be projected on graph

w_Year_Tracker #run the year tracker
anim_save("w_Year_Tracker.gif") #save the animation as a gif
```

Men's 100M Medal Times from 1896-2016
```{r}
m_Year_Tracker <- d_100M_Men %>% #pipe the filtered data through
  ggplot() + 
    geom_point(aes(x = Year, y = Result, col = Medal), 
    alpha = 0.8) + #set the parameter data for the plot
    theme(legend.position = "right") + guides(size = "none") + #create & set position of legend
    labs(x = "Year", y = "Run Time Results (sec)", title = "Men's 100M Medal Times (1896 - 2016)", col = "Medal Type") + #set the x and y axis labels and title 
    geom_text(aes(x = min(Year), y = min(Result), label = as.factor(Year)) , hjust=-2, vjust = -0.2, alpha = 0.2,  col = "gray", size = 20) + #use data to set column/row text and length of the graph
    transition_states(as.factor(Year), state_length = 0) #separates the data by year to be projected on graph

m_Year_Tracker #run the year tracker
anim_save("m_Year_Tracker.gif") #save the animation as a gif
```

**Nationality Times Race (Bar Race)**
This set of bar graphs plots the result times by Nationality of each medalled athlete.This section took me a while to figure out how to order the Medals (G, S, B) and correspond them to (1,2,3). But I GOT IT!!!\
[*Disclaimer* - It won't be much of a race since the times are so close, but none the less it's a RACE!!]\

Women's Race
```{r}
#create a new data frame that groups the data by year
w_race <- d_100M_Women %>%
  group_by(Year) %>%
  arrange(factor(Medal, levels = c('G', 'S', 'B')))%>% #arrange the Medal column for the order G,S,B
  arrange(Year)%>% #arrange the data by year
  mutate(ranking = row_number()) #create a ranking column to align with the medal type (G = 1, S = 2, B = 3)
  
head(w_race) #print/check the new data set  
```

```{r}
w_bar_Race <- w_race %>% #pipe the new dataset through
  ggplot() +
  geom_col(aes(ranking, Year, fill = Nationality)) +
  geom_text(aes(ranking, Year, label = Result), hjust = -0.1) +
  geom_text(aes(ranking, y = 0 , label = Nationality), hjust = 1.1) +
  geom_text(aes(x = 15, y = max(Year) , label = as.factor(Year)), vjust = 0.2, alpha = 0.5,  col = "gray", size = 20) +
  coord_flip(clip = "off", expand = FALSE) + scale_x_reverse() +
  theme_minimal() + theme(
    panel.grid = element_blank(), 
    legend.position = "none",
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    plot.margin = margin(1, 4, 1, 3, "cm")
  ) +
  transition_states(Year, state_length = 0, transition_length = 2) +
  enter_fade() + #fade the bar in 
  exit_fade() + #fade the bar out
  ease_aes('quadratic-in-out') #transition of the 2 components in and out 

wb_race <- animate(w_bar_Race, fps = 25, duration = 15, rewind = FALSE) #run the animation
wb_race
anim_save("wb_race.gif") #save the animation as a gif
```

Men's Race
```{r}
#create a new data frame that groups the data by year
m_race <- d_100M_Men %>%
  group_by(Year) %>%
  arrange(factor(Medal, levels = c('G', 'S', 'B')))%>% #arrange the Medal column for the order G,S,B
  arrange(Year)%>% #arrange the data by year
  mutate(ranking = row_number()) #create a ranking column to align with the medal type (G = 1, S = 2, B = 3)
  
head(m_race) #print/check the new data set 
```

```{r}
m_bar_Race <- m_race %>% #pipe the new dataset through
  ggplot() +
  geom_col(aes(ranking, Year, fill = Nationality)) +
  geom_text(aes(ranking, Year, label = Result), hjust = -0.1) +
  geom_text(aes(ranking, y = 0 , label = Nationality), hjust = 1.1) +
  geom_text(aes(x = 15, y = max(Year) , label = as.factor(Year)), vjust = 0.2, alpha = 0.5,  col = "gray", size = 20) +
  coord_flip(clip = "off", expand = FALSE) + scale_x_reverse() +
  theme_minimal() + theme(
    panel.grid = element_blank(), 
    legend.position = "none",
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    plot.margin = margin(1, 4, 1, 3, "cm")
  ) +
  transition_states(Year, state_length = 0, transition_length = 2) +
  enter_fade() + #fade the bar in 
  exit_fade() + #fade the bar out
  ease_aes('quadratic-in-out') #transition of the 2 components in and out 

mb_race <- animate(m_bar_Race, fps = 25, duration = 15, rewind = FALSE) #run the animation
mb_race
anim_save("mb_race.gif") #save the animation as a gif
```

